---
title: "Aperture - Git log data"
date: "`r Sys.Date()`"
output: github_document
---


```{r setup, echo = FALSE, cache = FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(rmdformats)
require(tidyverse)
library(plotly)
require(janitor)
require(skimr)
library(mosaic)
library(inspectdf)
library(visdat)
library(DT)
library(hrbrthemes)
# knitr chunk options ----
knitr::opts_chunk$set(
  echo = TRUE, # show/hide all code
  # results = "hide", # hide results
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  fig.width = 9, # figure width
  fig.height = 7, # figure height
  warning = FALSE, # show warnings
  size = "small", # size of the text
  fig.path = "figs/"
) # location of files
# knit settings ----
knitr::opts_knit$set(
  width = 78
)
# base options ----
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000,
  max.print = 999999
)
```

```{r packages, message=FALSE, warning=FALSE}
library(cloc) # cloc (counting lines of code)
library(gh) # github api package
library(tidyverse) # tidy analysis tools
library(tidymodels) # tidy modeling
library(tidygraph) # tidy graphs
library(tidytext) # text analysis for commit messages 
library(ggraph) # graphing
library(janitor) # cleaning data
library(inspectdf) # visualizing entire data sets 
library(skimr) # summary stats 
library(visdat) # missing data
```

Check out the repository for this project [here]().

# Motivation

This document covers how to use the `git log` to get meaningful metrics about your workflow and code. Great resources for learning how to use the `git log` tools are, 

1) the [online git documentation](https://git-scm.com/doc), 
2) [Professional Git](https://www.amazon.com/Professional-Git-Brent-Laster/dp/111928497X),  
3) [Pro Git](https://git-scm.com/book/en/v2) and  
4) [Git for Teams](http://gitforteams.com/) 

## An example repo

In order to replicate the work in the Appendix 3 of Software Design X-Rays (which covers author summaries, `cloc`, and complexity), we'll need a repo. I've chosen the `ggplot2` repository because it's fairly common (and I know it's widely used by data scientists).

Below is a code chunk to clone the `ggplot2` repo from Github into the parent folder. 

```{bash, clone-ggplot2-repo, eval = FALSE}
cd downloads
pwd
git clone https://github.com/tidyverse/ggplot2.git
```

Let's check the contents of this downloaded folder. 

```{r ggplto2-tree-structure}
fs::dir_tree("downloads/ggplot2", recurse = FALSE)
```

## Folder structures 

For the sake of reproduction, the folder structure for this project is below. 

```{r parent-folder}
fs::dir_tree(".", recurse = FALSE)
```

### Path to `ggplot2` folder 

The path to the `ggplot2` package folder is here: 

```{r ggplot2_folder_path}
ggplot2_folder_path <- here::here("downloads", "ggplot2")
ggplot2_folder_path
```

We can run `bash` scritps in RStudio, so let's walk through the `regex` pattern.

From `SDX`(page 8):

> We use the same mental processes to understand code as those we use in everyday life beyond our keyboards (evolution wasn’t kind enough to equip our brains with a coding center). 

> As we learn a topic we build mental representations of that domain. Psychologists refer to such mental models as schemas. A schema is a theoretical construct used to describe the way we organize knowledge in our memory and how we use that knowledge for a particular event. **You can think of a schema as a mental script implemented in neurons rather than code.**

This is a handy way to think about code. 

## Command line tools/regex for Hotspots

> The command then counts the frequency of the resulting file names and delivers the results sorted in descending order:

The regex pattern below is using the pipe `|` operator to chain together the git and unix commands. 

### Extracting `git log` data

The `git log` command returns a log of all commits. We add three options: 
1. [`--format=format:`](https://git-scm.com/docs/git-log#_commit_formatting). This tells git how to return the logs.

2. [`--name-only`](https://git-scm.com/docs/git-log#Documentation/git-log.txt---name-only). This is the name of the changed files.

3. `head -n 20` - We will look at the top 20 lines from this command.

```{bash, format-name-only}
# change working directory
cd downloads/ggplot2 
# verify
pwd 
# /Users/martinfrigaard/@Working/tech-debt/ggplot2
# now run the regex with cloc
git log --format=format: --name-only | head -n 20
```

There are too many lines and whitespace between each output, so it needs some wrangling.

### egrep 

The `egrep` command removes all the lines in between each commit log item. Read more [here](https://earthsci.stanford.edu/computing/unix/editing/grep.php).  

The added `-v` tells bash the "selected lines are those **not** matching any of the specified patterns."

```{bash, egrep}
# change dir
cd downloads/ggplot2 
git log --format=format: --name-only | egrep -v '^$' | head -n 10
```

This cleans the output up and removes some of the whitespace. 

### sort and unique

The next few commands sorts the output from the Git log, and then collapses them to each unique entry.

1. The first [`sort`](https://ss64.com/bash/sort.html) arranges everything alphabetically, or *Sort text files. Sort, merge, or compare all the lines from the files given (or standard input.)*. 

2. The `uniq -c` command reduces each file to the [unique name and count](https://ss64.com/bash/uniq.html). 

3. The second [`sort -r`](https://ss64.com/bash/sort.html) puts the highest numbers to the top, or *reverse the result of comparison, so that lines with greater key values appear earlier in the output instead of later*.

```{bash sort}
# change dir
cd downloads/ggplot2 
# test the next
git log --format=format: --name-only | egrep -v '^$' | sort | uniq -c | head -n 10
```

Now these contents are sorted.

```{bash sort-uniq}
# change dir
cd downloads/ggplot2 
# test the next
git log --format=format: --name-only | egrep -v '^$' | sort | uniq -c | head -n 20
```

I can ignore the `sort: Broken pipe` message, because I am only viewing the output here. So now we know to get the output from the `git log`. We can create a new file called `file_name_counts.txt` and send the results to this file. 

```{bash file_name_counts.txt}
# change dir
cd downloads/ggplot2 
# create new file
touch "file_name_counts.txt"
# verify 
ls
```

We created the file, now we have to put the output into it with the next chunk.

```{bash pipe-to-file_name_counts.txt}
# change dir
cd downloads/ggplot2 
# repeat commands and send to new file
git log --format=format: --name-only | egrep -v '^$' | sort | uniq -c | sort -r >> "file_name_counts.txt"
```

Now we will move this file into current directory and read them into R. 

```{r move-file_name_counts.txt}
# check current location
# fs::dir_ls("../ggplot2")
fs::file_move(path = "downloads/ggplot2/file_name_counts.txt", 
              new_path = paste0("data/", 
                                base::noquote(lubridate::today()), 
                                "-file_name_counts.txt"))
# verify new location
fs::dir_ls("data")
```

Now we can load these data into R. 

```{r export-GitFileCount}
GitFileCount <- readr::read_delim(file = paste0("data/", 
                                            base::noquote(lubridate::today()), 
                                            "-file_name_counts.txt"),
                                         delim = " ", 
                                         skip = 1, 
                                  # column names
                                         col_names = c("file_changes", 
                                                       "gitlog_path"))
```

Now we have a data fame with the files and their total number of changes. 

```{r skim}
GitFileCount %>% skimr::skim()
```


These data are showing the `gitlog_path`, so we will create a new variable with only the `filename`.

```{r create-GitFileCount-filename}
GitFileCount <- GitFileCount %>% 
  dplyr::mutate(filename = base::basename(gitlog_path))
GitFileCount$filename %>% utils::head()
```

We'll make some changes to the variables so they're easier to plot. 

```{r mutate}
GitFileCount <- GitFileCount %>% 
  dplyr::mutate(file_changes = as.numeric(file_changes),
                filename = as.factor(filename)) 
GitFileCount %>% dplyr::glimpse(78)
```

# Visualize file changes 

Below we set a theme for some better looking graphics.

```{r set-theme}
ggplot2::theme_set(hrbrthemes::theme_ipsum_tw(
  base_size = 10,
  strip_text_size = 11,
  axis_title_size = 13,
  plot_title_size = 17,
  subtitle_size = 13,
  base_family = "EnvyCodeR",
  # "JosefinSans-LightItalic"
  strip_text_family = "TitilliumWeb-Regular",
  axis_title_family = "TitilliumWeb-Regular",
  subtitle_family = "TitilliumWeb-Regular",
  plot_title_family = "JosefinSans-Regular"
))
```


The graph below is a Cleveland dot plot of the files that had more than 75 changes.

```{r graph-file-counts}
# define labs first!
file_count_labs <- ggplot2::labs(
    y = "Number of file changes",
    x = "Name of file",
    title = "File changes from git log",
    subtitle = "ggplot2 files with > 75 changes",
    caption = "from git log --format=format: --name-only") 
# now plot these data in a scatter plot
GitFileCount %>% 
  dplyr::filter(file_changes > 75) %>%  
  ggplot2::ggplot(aes(x = fct_reorder(.f = filename,
                             .x = file_changes),
             y = file_changes)) +
  ggplot2::geom_point() + 
  ggplot2::coord_flip() +
  file_count_labs
```

This makes sense, because the `NEWS`, `DESCRIPTION`, and `NAMESPACE` are the most commonly touched files. 

## Next steps

In the next section (`02-counting-lines-of-code.Rmd`), we will be counting the lines of code in each file.

```{r export-file-data}
# ls()
readr::write_csv(as.data.frame(GitFileCount), 
                 path = paste0("data/", 
                         base::noquote(lubridate::today()), 
                           "-GitFileCount.csv"))
fs::dir_tree("data")
```

