---
title: "Part 04 gitsum: A package for exploring Github data"
output: github_document
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)
require(tidyverse)
library(plotly)
require(janitor)
require(skimr)
library(gh)
library(mosaic)
library(inspectdf)
library(visdat)
library(DT)
library(formatR)
library(hrbrthemes)
library(gcookbook)
library(stringi)
library(ggbeeswarm)
library(listviewer)
library(hrbrthemes)
## Global options
options(max.print = "75")
# knitr chunk options ----
knitr::opts_chunk$set(
  echo = TRUE, # show/hide all code
  tidy = FALSE, # cleaner code printing
  comment = "#> ", # better console printing
  eval = TRUE, # turn this to FALSE stop code chunks from running
  message = TRUE, # show messages
  fig.width = 9, # figure width
  fig.height = 6, # figure height
  warning = FALSE, # show warnings
  size = "small", # size of the text
  fig.path = "figs/"
) # location of files
# knit settings ----
knitr::opts_knit$set(
  width = 78
)
# base options ----
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000,
  max.print = 999999
)
ggplot2::theme_set(hrbrthemes::theme_ipsum_ps(
  base_size = 9,
  strip_text_size = 10,
  axis_title_size = 11,
  plot_title_size = 15,
  subtitle_size = 11,
  base_family = "AndaleMono",
  strip_text_family = "TitilliumWeb-Regular",
  axis_title_family = "TitilliumWeb-Regular",
  subtitle_family = "TitilliumWeb-Regular",
  plot_title_family = "JosefinSans-Regular"
))
```

## Install

This is the install of the latest version from [Github here](https://lorenzwalthert.github.io/gitsum/). 

```{r install}
devtools::install_github("lorenzwalthert/gitsum")
```

## Load

Load the package with `library()`.

```{r load}
library(gitsum)
```

## Overview 

[`gitsum`](https://lorenzwalthert.github.io/gitsum/) is a package from [Lorenz Walthert](https://github.com/lorenzwalthert) that's designed to "*parse a git repository history to collect comprehensive information about the activity in the repo. The parsed data is made available to the user in a tabular format. The package can also generate reports based on the parse data.*

## Functions in `gitsum`

The documentation on `gitsum` can be found in the Github repo for this package. We can find the following files from the `downloads/2020-02-17-gitsum` folder. 

```{r gitsum-R}
fs::dir_tree("downloads/2020-02-17-gitsum", recurse = FALSE)
```

## `gitsum` - an example

To demonstrate the abilities of `gitusm`, we're going to look at RStudio's [`ggplot2`](https://ggplot2.tidyverse.org/), an open source data visualization tool. 

The package has been downloaded into the `downloads` folder here: 

```{r downloads}
fs::dir_tree("downloads/ggplot2", recurse = FALSE)
```

We will start with the `parse_log_simple()` and put the file path to the `downloads/ggplot2` folder. 

```{r parse_log_simple}
Ggp2PLogSimple <- gitsum::parse_log_simple(path = "downloads/ggplot2")
Ggp2PLogSimple %>% utils::head(10)
```


There are two main functions for parsing the history, both return tabular data:

- `parse_log_simple()` is a relatively fast parser and returns a tibble with one commit per row. There is no file-specific information.



`parse_log_detailed()` outputs a nested tibble and for each commit, the names of the amended files, number of lines changed ect. available. This function is slower.




