---
title: "code maat - installing and running clojure"
author: "Martin Frigaard, MA, MAS"
date: "`r Sys.Date()`"
output: 
  github_document: 
    fig_height: 6
    fig_width: 8
    toc: yes
---

```{r setup, include=FALSE}
library(hrbrthemes)
library(extrafont)
library(knitr)
library(rmdformats)
library(tidyverse)
library(magrittr)
library(skimr)
library(rlang)
library(gt)
library(visdat)
library(inspectdf)
library(foreign)
library(naniar)
# figs folder
if (!file.exists("figs")) {
  dir.create("figs")
}
knitr::opts_chunk$set(
  eval = TRUE, # we will run these as needed!
  echo = TRUE,
  cache = FALSE,
  prompt = FALSE,
  tidy = FALSE,
  fig.width = 10,
  fig.height = 7.5,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.path = "figs/"
)
knitr::opts_knit$set(
  width = 78
)
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000
)
options(max.print = 10000)
# theme ----
ggplot2::theme_set(hrbrthemes::theme_ipsum_tw(
  base_size = 10,
  strip_text_size = 11,
  axis_title_size = 13,
  plot_title_size = 17,
  subtitle_size = 13,
  base_family = "Ubuntu",
  strip_text_family = "TitilliumWeb-Regular",
  axis_title_family = "TitilliumWeb-Regular",
  subtitle_family = "TitilliumWeb-Regular",
  plot_title_family = "JosefinSans-Regular"
))
```

# Installing code maat 

code maat tool is free and open source on [Github](https://github.com/adamtornhill/code-maat). This tool was developed by Tornhill and is implemented in Clojure. 

## Step 1: Print Java version 

The version of Java is accessible via `java -version`

```{bash}
java -version
```

The Java version on this machine is new enough.

## Step 2: `code maat` jar files

We've downloaded the jar files from Tornhill's website, available [here](http://adamtornhill.com/code/maatdistro.htm). These files are stored in the `downloads` folder. 

```{r downloads, eval=TRUE}
fs::dir_tree("downloads", recurse = FALSE)
```

## Generate git log file

```{bash}
git log --all --numstat --date=short --pretty=format:'--%h--%ad--%aN' --no-renames --after=2018-01-31 > logfile.log
# cat logfile.log
```


### Quick description of code maat

What does code maat do?

> Code Maat operates on the level of log files generated from any of several version-control systems (including Subversion, TFS, Git, and others), and the GitHub page contains detailed instructions on how to generate the input data to the tool. 

We'll go through the installation instructions on page 216 of the SDXR text. 

> Once you have a log file you feed it to Code Maat, which parses the log file and then performs any of the `file-` or `architectural-` level analyses discussed in this book. You specify the type of analysis via a command-line argument. Here’s an example of a change coupling analysis, specified through the `-a` coupling argument, and using one of the supported Git formats where the version-control log is stored in the file `vcs_log.txt`:

```bash
$ java -jar code-maat-standalone.jar -c git2 -l maat.log -a coupling

entity,coupled,degree,average-revs
analysis/effort.clj,analysis/effort_test.clj,100,5
analysis/churn.clj,analysis/churn_test.clj,89,15
parsers/git.clj,parsers/git_test.clj,80,24
```

> Since pure text is the universal interface it allows you to postprocess that data or visualize it as discussed later in this appendix.

Where is `vsc_log.txt`? We need a bit more information to get this setup, so we'll head over to the Github page. 

## Step 3: Install `leinigen` (for `uberjar`)

First we need to install `leinigen` package with homebrew: 

```{bash}
brew install leiningen
```

Read more about [homebrew packaging](https://github.com/technomancy/leiningen/wiki/Packaging).

## Step 4: using `uberjar` with leiningen

We'll begin by building the .jar code from source using `lein uberjar` [see source here](https://github.com/adamtornhill/code-maat#usage), but we need to make sure the clojure project file has been created and saved in the same directory we are currently working in.

### Step 4.1: Create a `project.clj` file

We need to put the following contents in a file titled, `project.clj`

```{clojure, eval=FALSE}
(defproject myproject "0.5.0-SNAPSHOT"
  :description "A project for doing things."
  :license "Eclipse Public License 1.0"
  :url "http://github.com/technomancy/myproject"
  :dependencies [[org.clojure/clojure "1.8.0"]]
  :plugins [[lein-tar "3.2.0"]])
```

This file is located in the parent directory, 

```{r project.clj}
fs::dir_tree(".", recurse = FALSE)
```

After installation, we can proceed with the installation of the `uberjar` package.  

```{bash}
lein uberjar
```

Executing the code file below creates the `target/` directory.

```{r target}
fs::dir_tree("target")
```

## Step 5: Creating a `logfile.log`

In order to store the output data in a file we can access, we'll need to create a log file using `touch` and the name of the file, like `logfile.log`.

```{bash}
touch "logfile.log"
# verify
ls
```

We can see the `logfile.log` file in the parent folder. 

## Step 6: Running code maat 


We can now use the `java -jar` command coupled with `code-maat-1.0-standalone.jar` to run `code maat` with the `bash` commands below.

```{bash}
java -jar code-maat-1.0-standalone.jar -l logfile.log -c git2 > code_maat_log.csv
# cat code_maat_log.csv
```

This shows us the `entity,n-authors,n-revs`. 

### Print `-summary`

Now that everything seems to be working, we can use the `-c git2 -a summary` commands to get the commits, entities, entities-changed, and authors. 

```{bash}
java -jar code-maat-1.0-standalone.jar -l logfile.log -c git2 -a summary
```

### Running code maat in `code-maat` repo (super meta)

We will run the same code from above in the [`code maat` package repo](https://github.com/adamtornhill/code-maat) to see if it works on an outside Git repo. 

```{bash}
cd downloads/code-maat-master
# check to see where you are!
# pwd
# create logfile.log
git log --all --numstat --date=short --pretty=format:'--%h--%ad--%aN' --no-renames --after=2017-01-31 > logfile.log
# verify new file
# ls -la 
```

```{bash}
cd downloads/code-maat-master
## ===== run the git2 java jar script ===== ##
# /-- this is the previous script:
#     java -jar code-maat-standalone.jar -c git2 -l maat.log -a coupling 
# --/ we'll be repeating this but with the new standalone.jar
java -jar code-maat-1.0-standalone.jar -l logfile.log -c git2 > code_maat_master_log.csv
```

Now we can check the code maat file result. 

```{bash}
cd downloads/code-maat-master
head -n 7 code_maat_master_log.csv
```

If we import this into RStudio, we can visualize the 

### Print `-a coupling`

Now we will create another log file (couplinglog)

```{bash}
# cd downloads/code-maat-master
git log --all --numstat --date=short --pretty=format:'--%h--%ad--%aN' --no-renames --after=2017-01-31 > couplinglog.log
java -jar code-maat-1.0-standalone.jar -l logfile.log -c git2 -a coupling
```

This is showing the headers for the data, but not the actual numbers...

>  These repos aren't showing any data--any ideas?

```{bash}

```


